---
output:
  pdf_document:
    toc: no
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
  bookdown::html_document2:
    fig_caption: yes
    toc: no
  word_document: default
bibliography:
header-includes:  \usepackage{lineno} \usepackage{placeins} \usepackage{setspace}\doublespacing
# editor_options: 
#   chunk_output_type: console
---

\begin{center}
	
\textbf{\Large PUT YOUR TITLE HERE}
	
\textsc{AUTHOR1$^{1*}$, AUTHOR2$^{2}$, AUTHOR3$^{3}$}
\vspace{3 mm}
\normalsize{\indent $^1$AFFILLIATION 1 \\ $^2$AFFILIATION2 \\ $^3$AFFILIATION3}
$\text{*}$ Corresponding authors: CORRESPONDING AUTHOR NAME AND CONTACT INFO (IN THE AUTHOR SECTION, PUT THE ASTRIK NEXT TO THE CORRESPONDING AUTHOR)
\end{center}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev="cairo_pdf", cache = FALSE)

# THIS IS YOUR SET UP CHUNK. HERE YOU UPLOAD PACKAGES AND DATA
```

\newpage

\linenumbers

# ABSTRACT

WRITE YOUR ABSTRACT HERE

 Keywords: WRITE KEYWORDS HERE

# THIS IS HOW YOU MAKE A NEW HEADER

## THIS IS HOW YOU MAKE A SUB HEADER

THIS IS HOW YOU WRITE NORMAL TEXT (NO # SIGN). THIS IS WHERE YOU DO YOUR NORMAL WRITING. YOU CAN REFERENCE FIGURES BY \@ref(fig:NAMECODECHUNK)


<!-- IN ORDER TO MAKE A REFERENCE, YOU MUST WRITE [@BibtexCODEFORREFERENCE].  -->

```{r NAMECODECHUNK, fig.cap = 'CHANGE FIGURE CAPTION HERE'}

plot(x = mtcars$mpg, y = mtcars$qsec)

#THIS IS CALLED A CODE CHUNK, WHERE YOU WRITE R, TO MAKE A NEW CODE CHUNK, COPY THIS CODE WITHOUT THE COMMENT. EACH CODE CHUNK MUST HAVE A DIFFERENT NAME

# ```{r CHANGECODECHUNKNAME, fig.cap = "CHANGE FIG CAP"}
#   #WRITE R CODE HERE
# 
# ```



```
