---
title: "R Tutorial"
author: "Sophie Wulfing"
date: "2024-01-20"
output: html_document
---
# R STUDIO

## 4 PANELS
### UPPER LEFT - SOURCE PANEL
- Write and save code scripts to recreate analyses

### LOWER LEFT - CONSOLE
- Rusults of code. Can also run code if you want to check results or interact with your date

### UPPER RIGHT
- Environment has variables you've created

### LOWER RIGHT
- Tabs: Files - Can explore your R files
  Plots: Visulizations
  Packages: Where you can download programs other people have created to run specific functions 
  Help: access library SHOW EXAMPLE OF ?

# CREATING A DOCUMENT AND START CODING
- File => New file => R script
- Ways to run code
  Highlight, Code Run
  Run in upper right
  CTRL + ENTER
  
``` {r BASIC OPERATORS}
#Operator dasar
2 + 3

#Division
94729 / 7

# Integer Division
94729 %/% 7

#Modulus - Remainder
94729 %% 7

#YOU CAN COMMENT OUT CODE LIKE THIS

#PRINT FUNCTION

print("Hello, world")


```
# VARIABLES
Method of storing data

 - MUST start with letter
 - Letters/digits/periods
 - Case sensitive
 - Unacceptable variable names: TRUE, FALSE, NULL, if. Already functions in R 
 
 <!-- karena itu adalah fungsi -->
 
```{r Variables}


# 1name <- "Sophie"
# .1name <- "Sophie"
# _1name <- "Sophie"
# name 1 <- "Sophie"

#Variables are case-sensitve
name1 <- "Sophie"
NAME1

x <- 4
y <- 17

x*y

z <- x * y
z

#LATIHAN 1

#Concatinate Strings
Country <- "Indonesia"
Age <- 27

paste("My name is", name1, "I am from", Country, "I am", Age, "Years Old") #Need commas between



```
# DATA TYPES
- Numeric
- Integer/complex
- Boolean/logical
- Character
- String
- Factor - Skip for now, strings with assigned hierarchy

```{r dataTypes}

# NUMERIC 
k <- "17"
x * k

class(k)
k <- as.numeric(k)


class(k)
x * k

# BOOLEAN: TRUE OR FALSE. See handout
x < k
x == k #need two ==


# Characters and strings
nchar(name1)
grepl("Soph", name1)
grepl("Anne", name1)

```
# LISTS
An ordered collection of objects
Can store multiple objects in one variable


```{r Lists}

# GENERATING LISTS

fruit <- c("apple", "banana", "watermelon", "snake fruit", "mango")
numbersList <- c(4, 546, 234, 85, 1, 8, 3543, 654, 14, 46, 8934)

OneToTen <- 1:10
evenNumbers <- seq(0, 20, 2)
tenOnes <- rep(3, 5)

# INDEXING LISTS
min(numbersList)
max(numbersList)
sort(numbersList)
length(numbersList)
range(numbersList)

fruit[1]
fruit[1:3]
fruit[c(1,3, 5)]
fruit[-2]

fruit[1] <- "jeruk"
fruit

#boolean operator
"apple" %in% fruit

#Append
fruit_2 <- append(fruit, "apple", after = 2)
fruit_2

```
 # MATRICES, DATA FRAMES, AND NESTED LISTS
 
 Two or higher dimensional datasets with rows and columns
 
```{r matrices}

# Create a matrix
numbersMatrix <- matrix(c(1,2,3,4,5,6), nrow = 2, ncol = 3)

# Print the matrix
numbersMatrix

# Can be done with strings as well
fruitMatrix <- matrix(c("apple", "banana", "cherry", "mango","grape", "pineapple", "pear", "melon", "fig"), nrow = 3, ncol = 3)

fruitMatrix

# Indexing
fruitMatrix[1, 2]
fruitMatrix[2,]
fruitMatrix[,2]

#More than one row or column
fruitMatrix[c(1,2),]
fruitMatrix[,c(1,2)]


# ADDING COLUMNS
newFruitMatrix_1 <- cbind(fruitMatrix, c("strawberry", "blueberry", "raspberry"))

# Print the new matrix
newFruitMatrix_1

#ADDING ROWS
newFruitMatrix_2 <- rbind(fruitMatrix, c("strawberry", "blueberry", "raspberry"))

# Print the new matrix
newFruitMatrix_2

# REMOVING ROWS AND COLUMNS
fruitMatrix <- fruitMatrix[-c(1), -c(1)]

fruitMatrix

# check if in matrix
"apple" %in% fruitMatrix
"grape" %in% fruitMatrix

# Check Sizes
dim(fruitMatrix)
length(fruitMatrix)

```

```{r dataframes}

# Create a data frame
Data_Frame <- data.frame (
  Student = c("Rina", "Virginia", "Ben"),
  Score = c(100, 150, 120),
  Time = c(60, 30, 45)
)

# Print the data frame
View(Data_Frame)

summary(Data_Frame)

#Accessing
View(Data_Frame[1])

Data_Frame[["Student"]]

Data_Frame$Student


# Add a new row
New_row_DF <- rbind(Data_Frame, c("John", 110, 110))

# Print the new row
View(New_row_DF)

# Add a new column
New_col_DF <- cbind(Data_Frame, Place = c(2, 1, 3))

# Print the new column
View(New_col_DF)

# Remove the first row and column
Data_Frame_Remove <- Data_Frame[-c(1), -c(1)]

# Print the new data frame
View(Data_Frame_Remove)

# Dimensions
dim(Data_Frame)
ncol(Data_Frame)
nrow(Data_Frame)
length(Data_Frame) #same as ncol
```
 
 # IF STATEMENTS
 
 Will execute code only if a statement is true
 
```{r ifstatement}
#if statement
a <- 33
b <- 200

if (b > a) { #Boolean statement. If true, will execute what's in the curly brackets
  print("b is greater than a")
}

# else if statement - run if the previous statement is untrue

a <- 33
b <- 33

if (b > a) {
  print("b is greater than a")
} else if (a == b) {
  print ("a and b are equal")
}

# Else statement - if no previous statements are true

a <- 200
b <- 33

if (b > a) {
  print("b is greater than a")
} else if (a == b) {
  print("a and b are equal")
} else {
  print("a is greater than b")
}

```

``` {r nestedifstatement}

# Else statement SHOW RUSSIAN DOLL EXAMPLE

a <- 200
b <- 33

if (a < 1000){
  if(b < 1000) {
  print("both a and b are below 1000")
  } 
}



```


``` {r qualifstatement}

# 2 statement

a <- 200
b <- 33

if (a < 100 & b < 100) {
  print("both a and b are below 100")
} 

if (a < 100 | b < 100) {
  print("either a or b are below 100")
} 


```

# WHILE LOOPS

Will execute a code as long as a condition is true

```{r whileloops}

i <- 1
while (i <= 6) {
  print(i)
  i <- i + 1
}

```
 
# FOR LOOPS

Iterates over a sequence

```{r forLoops}
for (x in 1:10) {
  print(x)
}


fruits <- list("apple", "banana", "cherry")

for (x in fruits) {
  print(x)
}

```


# FUNCTIONS

A block of code that is run only when you call it

```{r functions}

my_function <- function() { # create a function with the name my_function
  print("Hello World!")
}

# Have to call the function to use it
my_function()

# Function with arguments

name_function <- function(name){
  print(paste("Hello, my name is", name))
}

name_function()
name_function("Sophie") #Now can change argument to change the output

# Multiple arguments
intro_function <- function(name, home){
  print(paste("Hello, my name is", name, "I am from", home))
}

intro_function("Sophie", "Washington")


# Function with return values - return is used more often because it stores a variable
mult_function <- function(x) {
  return (5 * x)
}

print(mult_function(3))
print(mult_function(5))
print(mult_function(9))

FinalScore <- mult_function(9) * 100
FinalScore
```
 
 
 
 
 
 
 
 
 
 
 
 
 
 